---
title: "Mid-term project"
author: "Pramod Hegde & Sandesh Shresta"
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: cerulean
---

# Mid-term project GitHub repository URL:

Paste here the URL of the GitHub project (after going through the instruction steps)
https://github.com/Promode45/midproject_group4


# Introduction

Gross Primary Production(GPP) and soil respiration make the highest carbon fluxes from the terrestrial ecosystem to the atmosphere (Bond-Lamberty and Thomson, 2010). Due to the magnitude of the CO2 flux from soils to atmosphere and the huge reserve of potential C in soils, any increases in soil respiration due to man-made or environmental changes may increase atmospheric CO2 levels and exacerbate global warming. Soil respiration consists mainly of respiration from roots, mycorrhizal fungi and microbes. Microbial respiration of soil C is the highest of the three (Wang and Fang, 2009). 

Human activities have been significantly modifying biogeochemical cycles in many ways. Management practices like tillage and fertilization are one of the key factors contributing to increasing CO2 efflux from soils. Nitrogen is the most important component for plant growth, is often limiting in ecosystems, and could affect soil respiration by altering microbial biomass, their activity and root growth (Ding et al.,2010). Studies have found a sharp increase in CO2 efflux following tillage, mainly due to broken soil clods and aggregates which act as "gas tankers" (Fielder et al., 2015; Morell et al., 2012). Although, sharp increases of CO2 efflux are seen following tillage, the extent of the increase depends on cultivation type, Soil Organic Carbon (SOC), soil texture and structure. 

Soil temperature and soil water content are considered the main environmental controls of soil respiration, as these can affect both microbial and root growth. The indirect effects of these factors, nutrient availability and plant productivity can also affect CO2 efflux from soils. Rainfall events can result in changes in both soil temperature and water content, and has been part of numerous studies (Shi et al., 2011; Muhr and Borken, 2009; Morell et al., 2012). But there are very few studies, where the combined effects of management practices, soil temperature and soil moisture have been studied. 

In this experiment, we try to see the effects of common land management practices on soil respiration across different days following precipitation. 

# Hypothesis and objectives

We hypothesize that, soil respiration increases with fertilization and tillage. Additionally, we believe that plots under these treatments will have a higher respiration following precipation as compared to the control plots. 

The objectives of this study are (a) to assess the effects of treatments fertilization, tillage or their combined effects on soil CO2 flux over six different days, (b) to assess the effects of precipitation on soil CO2 flux. 

# Material and Methods

## Site description

The study was conducted at the Whitehall forest in the Piedmont regions Clarke and Ocoonee county in Athens, Georgia, USA. The area has a humid subtropical climate with a mean annual precipitation of 1267.3 mm, and average January and July temperatures of 6.3 degree C and 26.3 degree C respectively. Average elevation is approximately 200m above mean sea level. 

## In-situ field measurements

The gas samples were taken in late January and early February following two precipitation events from 25 treatment plots. Each plot is fitted with Polyvinyl chloride collars at a depth of about 5cm which are used to collect gas samples. We used a Licor LI-870 gas analyzer (LI-COR Inc., Lincoln, NE, USA) to measure the CO2 flux. The device is also equipped with a probe that simultaneously measures soil moisture, soil temperature and Electrical Conductivity(EC). The measurements were taken on six different days to see the effects of rain on soil respiration. 
The study area experienced rainfall on 28th of January 2024, and we started our soil respiration measurements from the 29th of January 2024 on days 1, 2, 3, 8,9 and 10. The repeated measure in this experiment "days after rain" are these measurements that were taken between the 29th of January 2024 and 10th of February 2024. It is worth mentioning that, there was a slight precipitation on day 7, but we did not consider this rain event as significantly affecting the experiment. 

## Study design

The study consists of a set of five completely randomized blocks (within-site spatial replicates), each with five, 5x5m plots receiving either long-term nutrient addition, a physical disturbance, nutrient addition combined with disturbance, short-term nutrient addition or no treatments. The treatments were done just once at the beginning of summer every year (mid June). The randomization of the study site was carried out using a R code given on the project website (https://nutnet.org/sites/default/files/2019-07/create-site-maps-dragnet-03-July-2019.R).


## Statistical analysis
  
Soil respiration for different treatments were analyzed using a repeated measured mixed-effect ANOVA. Model explanatory variables included treatment (as categorical variable), days after rain (repeated measure) and their interaction as fixed effects. Plot which was nested inside blocks was modeled as the random effect. Different error correlation structures like ARMA(1), exponential, Gaussian,linear etc were applied and the one with the lowest Akaike Information Criterion (AIC) was selected for further analysis. The repeated measures were of unequal time intervals, and hence the choice of correlation structures were limited. Soil moisture and temperature were also measured, since these two factors are known to have an effect on how microbes respire carbon dioxide. The two factors were considered as covariates, and hence accordingly an ANCOVA was also conducted. The assumptions of ANCOVA are as follows - 

1. Linear relationships - the effect of the categorical variable on the outcome varies linearly with the outcome.  
2. Homogeneity of slopes - The relationship between the dependent and independent variable should be consistent across all the treatment levels. Slopes of regression lines (representing the effect of covariates) should be parallel for each group.  
3. Uncorrelated covariates - The covariates must not be correlated to the error term.  
  
We ran the ANCOVA model, and checked for one of the assumptions, which were not satisfied. Hence, we went ahead with the mixed-effects ANOVA. We used an alpha level of 0.05. We used the linear mixed effect mdoel (lme) from the package nlme (Pinheiro and Bates, 2000). The linearity assumptions were checked using the QQ-plot, and a point graph to check for normality, outliers and homoscedascity. **The assumption of outliers was not met, since we found two residuals that were off the limit of 3 standard deviations. But, the raw data points were checked and it did not seem out of place, and hence we decided to not remove them from our analysis.** Apart from this, all assumptions were satisfied without any problems. 

# Results

# Loading Packages
In this chunk, packages will be loaded that will be used in the process of the analysis.

```{r }
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#|message: False
#|warning: False
library(tidyverse)
library(janitor)
library(stringr)
library(ggpubr)
library(broom)
library(rstatix)
library(lme4)
library(nlme)
library(emmeans)
library(multcomp)
```

# importing data
We will import data in the following chunk

```{r }
#| echo: false
data <- read.csv("../Data/soilresp_midterm.csv")
summary(data)
```

# Data wrangling
In this chunk, we will manipulate and prepare the data for further analysis, including cleaning column header name, creating a column for rep(block), converting variables to appropriate data types and reshaping the data into a more suitable format for analysis.

```{r }
data_w <- data[-26,] %>% 
  clean_names() %>% 
  mutate(treatment = factor(treatment)) %>% 
  
  mutate(
    rep = case_when
    (plot == c(1:5) ~ "1",
    plot == c(6:10) ~ "2",
      plot == c(11:15) ~ "3",
      plot == c(16:20) ~ "4", 
      plot == c(21:25) ~ "5"
     )
  ) %>% 
  mutate(rep = factor(rep)) %>% 
  pivot_longer(cols =r1_01_29_24:r6_02_07_24,
               values_to = "respiration",
               names_to = "x1")   %>%
   mutate(x1 = gsub("resp", "", x1)) %>% 
 mutate(date1 = str_extract(x1, "\\d{2}_\\d{2}_\\d{2}")
  ) %>% 
  mutate(date = as.Date(date1, "%m_%d_%y")) %>% 
  mutate(dar = case_when(
    date == "2024-01-29" ~ "01",
    date == "2024-01-30" ~ "02", 
    date == "2024-01-31" ~ "03",
    date == "2024-02-05" ~ "08", 
    date == "2024-02-06" ~ "09",
    date == "2024-02-07" ~ "10"
  )) %>% 
  mutate(fdar = factor(dar)) %>%
  mutate(dar = as.numeric(dar)) %>% 
  dplyr::select(c(-x1,-date1)) %>%
  mutate(plot=factor(plot))


summary(data_w)
```
Some references for the data, swc stands for Soil Water Content, st stands for Soil temperature, 1-6 being the different days of measurement, dar being Days After Rain, date is the date of measurement. 

# EDA - Exploratory Data Analysis

```{r}
data_w %>% 
  group_by(fdar) %>% 
  ggplot(aes(treatment,respiration, 
             color = treatment
             ))+
  theme(legend.position = "none")+
  geom_boxplot()
```
From the above graph, we can make an initial analysis saying that the plots with NPK treatments have in general a higher respiration value, especially the combined effects of NPK and disturbance,

```{r}
data_w %>% 
  group_by(treatment) %>% 
  ggplot(aes(fdar, respiration,
             color = fdar))+
  labs(x = "Days After Rain (DAR)")+
  geom_boxplot()+
  theme_classic()+
  theme(legend.position = "none")+
  scale_color_viridis_d()
```
Although not too different, we could still say that respiration values the day right after a rain were higher than the other days. This is also part of our hypothesis. 

Here, we will inspect the distribution of respiration across different level of days after rain (dar), segmented by treatment, which can provide insights into potential relationships and patterns in the data.

```{r}
data_w %>% 
  ggplot(aes(fdar,respiration))+
  geom_boxplot()+
 facet_wrap(~treatment)
  
```
# Ancova
One of the assumptions of the ancova model is that the dependent variable must be linearly related to the covariates (Leppink, 2010)

In the chunk below, we will reshape data and visualize the data to check the linearity of the covariates.

```{r}
data_w %>% 
   pivot_longer(cols = starts_with("sw"),
               values_to = "watercontent", 
               names_to = "x2"
               ) %>% 
  mutate(x2 = gsub("water", "", x2)) %>% 
  pivot_longer(cols = st1:st6,
               values_to = "soiltemp",
               names_to = "x3") %>% 
  mutate(x3 = gsub("temp", "", x3)) %>% 
  ggscatter(x= "watercontent", y ="respiration",
            color = "treatment", add = "reg.line")+
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~"), color = treatment)
    )
```

The linearity assumption seems to be not satisfied, and thus we move ahead with just the anova analysis.

#  default model
Here, we will setup the environment for model fitting and fit the default linear mixed-effects model.

```{r}
options(contrasts = c("contr.sum", "contr.poly"))

mix_mod <- lme(respiration ~ treatment*fdar,
                random = ~ 1 |rep/plot,
                data = data_w)
mix_mod
```

In the following two chunks, we will gain further insights into the fitted linear mixed-effects model, including assessing the significance of fixed effects through ANOVA and examining the autocorrelation structure of the residuals.

```{r}
Anova(mix_mod,type = 3) 

```

```{r}
ACF(mix_mod, resType = "n") %>% 
  plot(alpha = 0.01)
```
Although the default model will not be considered for model comparison, it is an easy way to start, and get a handle of things. There are six lags in the ACF plot above, which agrees with the number of repeated measures. But, it does seem that lag 1 is beyond our threshold requirement. 

#  autroregressive moving average(ARMA)
In this chunk I will fit linear mixed-effects model with an autoregreessive moving average (ARMA) correlation structure to the data and assesses the autocorrelation structure of the residuals.

```{r}
mod2_ma <- lme(respiration ~ treatment*fdar,
                random = ~1 |rep/plot,
               correlation = corARMA(p = 1, q = 1),
                data = data_w)
mod2_ma

```
The output of the model fitted with ARMA correlation structure looks okay 
```{r}
ACF(mod2_ma,resType = "n") %>% 
  plot(alpha = 0.01)
```
The ACF plot looks good, all the lags are within the threshold. Now, we compare this model with the default one
```{r}
anova(mix_mod,mod2_ma)
```
The AIC value for model 2 is lower than the default, which is a positive thing. But, BIC values do not represent the same. But, we will move on, and try to apply different correlation structures to see their effectiveness. 


#  exponential correlation structure
In this chunk, I will fit the linear mixed model to the data, incorporating an exponential correlation structure to account for the correlation between observations based on the different level of days after rain.

```{r}
mod3_exp <- lme(respiration ~ treatment*fdar,
                random = ~1 |rep/plot,
                correlation = corExp(form = ~dar),
                data = data_w)
mod3_exp
```

```{r}
ACF(mod3_exp, resType = "n") %>%
  plot(alpha = 0.01)
```
One of the lags are beyonf the threshold, which we might be able to see in the comparison values as well, with a higher AIC value 
```{r}
anova(mix_mod,mod2_ma,mod3_exp)
```
As expected from the ACF curve, we do see that the exponential correlation structure did not perform well (higher AIC)

# Gaussian 
Here, we will fit the linear mixed model to the data, incorporating a Gaussian correlation structure and assess the autocorrelation structure of the residuals.

```{r }
mod4_gaus <- lme(respiration ~ treatment*fdar,
                random = ~1 |rep/plot,
                correlation = corGaus(form = ~dar),
                data = data_w)
mod4_gaus
```

```{r}
ACF(mod4_gaus, resType = "n") %>%
  plot(alpha = 0.01)
```

```{r}
anova(mix_mod,mod2_ma,mod3_exp, mod4_gaus)
```
Still model 2 with ARMA is the best performing. 

# Linear
Here, we will fit the linear mixed model to the data, incorporating a Linear correlation structure and assess the autocorrelation structure of the residuals.

```{r }
mod5_lin <- lme(respiration ~ treatment*fdar,
                random = ~1 |rep/plot,
                correlation = corLin(form = ~dar),
                data = data_w)

mod5_lin
```


```{r}
ACF(mod5_lin, resType = "n") %>%
  plot(alpha = 0.01)
```
```{r}
anova(mix_mod,mod2_ma,mod3_exp, mod4_gaus,mod5_lin)
```

# Rational quadratic

We will fit the linear mixed model to the data, incorporating a Rational Quadratic correlation structure and assess the autocorrelation structure of the residuals.

```{r }
mod6_rq <- lme(respiration ~ treatment*fdar,
                random = ~1 |rep/plot,
                correlation = corRatio(form = ~dar),
                data = data_w)
mod6_rq
```

```{r}
ACF(mod6_rq, resType = "n") %>%
  plot(alpha = 0.01)
```
```{r}
anova(mix_mod,mod2_ma,mod3_exp, mod4_gaus,mod5_lin,mod6_rq)
```

# Spherical 
We will fit the linear mixed model to the data, incorporating a Spherical correlation structure and assess the autocorrelation structure of the residuals.

```{r}
mod7_sph <- lme(respiration ~ treatment*fdar,
                random = ~1 |rep/plot,
                correlation = corSpher(form = ~dar),
                data = data_w)
mod7_sph
```

```{r}
ACF(mod7_sph, resType = "n") %>%
  plot(alpha = 0.01)
```

# Final comparison
Here, we will run a ANOVA based on different model, allowing for the comparison of the models based on their AIC values.

```{r}
anova(mix_mod, mod2_ma, mod3_exp, mod4_gaus, mod5_lin, mod6_rq, mod7_sph) %>%
  as.data.frame() %>%
  rownames_to_column(var = "modelname") %>%
  janitor::clean_names() %>%
  dplyr::select(modelname, model, df, aic, bic) %>%
  arrange(aic)
```
From the above table, we were able to say that Auturegressive Moving Average (ARMA) performed the best, since it had the lowest AIC values. We will base our analysis on the same. 

# Residual extraction
Here, we will extract the model residuals and calculate standardized residuals and keep in tidy format. 

```{r}
library(broom.mixed)
mod2_ma_resid <- augment(mod2_ma) %>% 
  mutate(.stdresid = resid(mod2_ma,
                           type = "pearson",
                           scaled = T
                           ))
mod2_ma_resid
```

# Block random effect
Here, random effects from the linear mixed model (mod2_ma) will be extracted and visualized using Quantitle-quantile(QQ) plot to assess the normality of the random effect.

```{r}
ranef(mod2_ma)[[1]] %>% 
   ggplot(aes(sample=`(Intercept)`))+
  stat_qq(  shape = 21,
            fill = "purple", 
            size = 3,
            alpha = .7
  )+
  stat_qq_line()+
  labs(x = "Theoretical quantile",
       y = "Sample quantile")+
  theme_bw()

```

The points seem a little off, but nothing too worrying. 

# blocks:plots random effects

```{r}
ranef(mod2_ma)[[2]] %>% 
   ggplot(aes(sample=`(Intercept)`))+
  stat_qq(  shape = 21,
            fill = "purple", 
            size = 3,
            alpha = .7
  )+
  stat_qq_line()+
  labs(x = "Theoretical quantile",
       y = "Sample quantile")+
  theme_bw()
```

The tails are a little off, but this is to be expected in limited number of data points. 

# within groups errors
Here we will run a diagnostic plot to assess the goodness-of-fit of the linear mixed effects model, examining the standardized residuals against fitted values to identify potential outliers and assess the trend in the residuals. 

```{r}
ggplot(mod2_ma_resid, aes(x=.fitted, y=.stdresid))+
  geom_hline(yintercept = 0, color="red")+
  geom_point(shape = 21,
             fill = "purple", 
             size = 3,
             alpha = .7)+
  geom_smooth()+
  geom_hline(yintercept = c(-3,3), color = "red")+
  theme_bw()
  
```
We do see that thre are two residual outliers, but upon further examination of our raw data we decided to leave the outliers as is. The raw data points did not seem too much out of the ordinary, and hence we had to make a personal choice of retaining the outlier, even though it might skew our analysis a bit. 

# Normality assumption of residuals
Here, we will run QQ plot to visualize whether the standardized residuals follow normal distribution. 

```{r}
ggplot(mod2_ma_resid, aes(sample=.stdresid))+
  stat_qq(  shape = 21,
            fill = "purple", 
            size = 3,
            alpha = .7
  )+
  stat_qq_line()+
  labs(x = "Theoretical quantile",
       y = "Sample quantile")+
  theme_bw()
```
Although the tails are a bit off, it does not seem to be a problem. 

Here, we will run type III ANOVA table for the fixed effects in mod2_ma and observe the significant factor.

```{r}
Anova(mod2_ma, type = 3)
```
We see a significant interaction effects, and hence we go ahead with the comparison of means of all treatments crossed with the days after rain. 

# treatment x fdar Interaction
Here, we compare means between different treatments within each day of measurement. 

```{r}
rm_rcbd_cld_tfdar <- emmeans(mod2_ma, ~ treatment|fdar) %>%
  cld(reversed = T,
      Letters = letters,
      adjust = "none") %>%
  as.data.frame() %>%
  mutate(letter = trimws(.group))
  
  rm_rcbd_cld_tfdar
```

# Final plot 1 (treatment in dar)
In this chunk, we will see the effects of different treatments on each of the different days. Different letters signify difference. 

```{r}
ggplot(mapping = aes(fill = treatment))+
  geom_boxplot(data = data_w,
               aes(x =treatment, 
                   y = respiration),
               alpha = 0.8) +
  geom_jitter(data = data_w,
              aes(x= treatment,
                  y = respiration),
              shape = 21,
              size = 1,
              alpha = 0.6) +
  geom_label(data = rm_rcbd_cld_tfdar,
             aes(x= treatment,
                 y = emmean,
                 label = letter),
             fill = "white",
             size = 2,
             alpha = 0.9) +
 labs(x = "Land Management",
       y = bquote("Soil respiration ("*mu* "mol" * "m" ^-2 * "s"^-1 *")"))+
  scale_fill_viridis_d() +
  facet_grid(.~fdar) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none",
        axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))
  
```

Overall, on all days soil efflux values from NPK + disturbance plots were higher than the other treatment plots, especially on day 1 and day 10 after rain, where we see a statistically significant difference between the combined treatment and the other treatments. It is also interesting to note the variability in efflux values in NPK plots. 

In the chunk below, we compare means of different days after rain within each treatment plots. 

```{r}
rm_rcbd_cld_dartf <- emmeans(mod2_ma, ~ fdar|treatment) %>%
  cld(reversed = T,
      Letters = letters,
      adjust = "none") %>%
  as.data.frame() %>%
  mutate(letter = trimws(.group))
rm_rcbd_cld_dartf

```

# Final plot 2 (dar in treatment)
In this chunk, we see how soil respiration varies across different levels of the drying cycle("fdar") highlighting significant difference between levels within treatment groups will be visualized using boxplot.

```{r}
ggplot(mapping = aes(fill = fdar))+
  geom_boxplot(data = data_w,
               aes(x =fdar, 
                   y = respiration),
               alpha = 0.8) +
  geom_jitter(data = data_w,
              aes(x= fdar,
                  y = respiration),
              shape = 21,
              size = 1,
              alpha = 0.6) +
  geom_label(data = rm_rcbd_cld_dartf,
             aes(x= fdar,
                 y = emmean,
                 label = letter),
             fill = "white",
             size = 2,
             alpha = 0.9) +
 labs(x = "Days after rain",
       y = bquote("Soil respiration ("*mu* "mol" * "m" ^-2 * "s"^-1 *")"))+
  scale_fill_viridis_d() +
  facet_grid(.~treatment) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none")
  
```

From the above graph, we can say that soil respiration spikes a day after rain in plots treated with NPK + disturbance. This confirms our hypothesis that soils with combined treatments of NPK + disturbance will have a higher CO2 efflux following a rain event. 

## Conclusion 
We hypothesized that we would see an increase CO2 efflux, which does seem to be the case as seen in the above graphs. The plots treated with both fertilization and tillage are higher on all days, with statistical difference on two of the days (day 1 and day 10). Plots treated with just NPK have a huge variation in their values but they do not seem to have any significant differences from the control. The same goes with plots with just disturbance as a treatment. 
Our hypothesis that precipitation can be a factor to increased fluxes is true in the plots treated with NPK + disturbance, as we can see the day right after a rain even (day 1), we saw a statistically significant increase in CO2 fluxes. But these effects were not found on other treatments. So, we could say that the interaction of NPK + disturbance and precipitation has an effect on the respiration values. 

# Team work in GitHub

Whether you are working with your future-self or as duos, make sure to stage, commit, and push after finishing each of the sub-sections above. When committing, write commit messages that are short and descriptive (e.g., finished wrangling).

If you are working in duos, make sure to split the workload. I will check your collaborative work through the commit history, and if one student has contributed significantly more than the other, than that will impact grades.

**Tip 1**: to avoid merge conflicts, make sure to **pull** first, and then start working locally. That will ensure that any changes made by your partner will be "downloaded" before you make changes to the files locally.

**Tip 2**: make use of the Issues on this repository to set up to-do lists and assign tasks to different people. You can also use each issue/task to discuss how things should be run and get to an agreement.

# Submitting your work

Once you have developed all the code and answers, make sure to Render this quarto file.


Rename this file to `LASTNAME1-LASTNAME2_midtermproject.html`.\
For ex., `Bastos-Mendes_midtermproject.html`.

Send the html file to my email (lmbastos\@uga.edu) by **April 11th** 11:59 pm.

# References 
1.Bond-Lamberty, B. and Thomson, A. (2010). Temperature-associated increases in the global soil
respiration record. Nature, 464(7288):579–582.
2. Ding, W., Yu, H., Cai, Z., Han, F., & Xu, Z. (2010). Responses of soil respiration to N fertilization in a loamy soil under maize cultivation. Geoderma, 155(3-4), 381-389.
3.Fiedler, S. R., Buczko, U., Jurasinski, G., & Glatzel, S. (2015). Soil respiration after tillage under different fertiliser treatments–implications for modelling and balancing. Soil and Tillage Research, 150, 30-42.
4.Morell, F. J., Whitmore, A. P., Álvaro-Fuentes, J., Lampurlanés, J., & Cantero-Martínez, C. (2012). Root respiration of barley in a semiarid Mediterranean agroecosystem: field and modelling approaches. Plant and soil, 351, 135-147.
5.Muhr, J., & Borken, W. (2009). Delayed recovery of soil respiration after wetting of dry soil further reduces C losses from a Norway spruce forest soil. Journal of Geophysical Research: Biogeosciences , 114 (G4).
6. Leppink, J. (2018). Analysis of Covariance (ANCOVA) vs. Moderated Regression (MODREG): Why the Interaction Matters. Health Professions Education, 4(3), 225–232. https://doi.org/10.1016/j.hpe.2018.04.001.
7.Pinheiro J, Bates D, R Core Team (2023). nlme: Linear and Nonlinear Mixed Effects Models. R package version 3.1-164, https://CRAN.R-project.org/package=nlme.
8.Shi, W. Y., Tateno, R., Zhang, J. G., Wang, Y. L., Yamanaka, N., & Du, S. (2011). Response of soil respiration to precipitation during the dry season in two typical forest stands in the forest–grassland transition zone of the Loess Plateau. Agricultural and forest meteorology, 151(7), 854-863.
9.Wang, W., & Fang, J. (2009). Soil respiration and human effects on global grasslands. Global and Planetary Change , 67 (1-2), 20-28.
